Decorrelating Errors in Quantum Gates by Random Gate Synthesis

### 2018-09-20 decorrelation/decorrelation.tex
RB Plots Updated
some changes to the text

\newcommand{\erf}[1]{Eq.~(\ref{#1})}
\newcommand{\needcite}{{\color{blue}\textsuperscript{[citation needed]}}}
\newcommand{\note}[1]{{\color{red}[#1]}}
\newcommand{\kcy}[1]{{\color{red}[#1]_{\rm{KCY}}}}

%-------------Header begins here----------------------------------------
\begin{document}
% \tableofcontents
\title{Decorrelating Errors in Quantum Gates by Random Gate Synthesis}

@@

\begin{abstract}
Thresholds for fault-tolerant quantum computation are often calculated assuming a noise model in which errors are uncorrelated. While convenient for simulation, these error models are often unphysical. Work by Preskill and others has shown that arbitrarily long computations may be performed even in the presence of spatial correlation, provided the correlation is sufficiently weak and decays sufficiently quickly with distance, but at the cost of a significantly lower threshold. The success of algebraic decorrelation methods, such as dynamical decoupling, demonstrate that quantum control techniques are capable of reducing temporal noise correlations. We propose to introduce similar methods to effect the spatial decorrelation of errors in quantum circuits, thereby increasing the threshold for fault-tolerant computation in such systems.
\end{abstract}

@@

Steady progress has been made in the theory of quantum error correction, proving higher thresholds for increasingly general models of noise \cite{Aharonov2006} \kcy{more recent citations}. These results show that quantum computation is feasible in principle, however recent NISQ \cite{Preskill2018} devices have noise that is not only often above thresholds, but that also violates fundamental assumptions made by the models used in these results, such as Markovianity \cite{Kitaev1997} and independence of errors\needcite. With these assumptions violated, promises about system performance and correctness are difficult to make.

There are many ways in which device noise can fail to be Markovian,\cite{Kelly2018, BlumeKohout2017}, and even when the noise \textit{is} Markovian, many models make simplifying assumptions about the structure of the noise. For instance, Pauli channels are often used to model systems\needcite due to their classical simulability\needcite, despite their unrealistic form. Some authors have then taken the approach of using these easily-understood models to upper bound the error in a real system by \textit{honestly} representing them \cite{Puzzuoli2014}. This approach is correct and rigorous, but produces overly-pessimistic bounds on device performance.

Other authors have addressed these problems \cite{Wallman2016, Campbell2017, Heim2016} by using circuit compilation to remove correlated noise. Given a process that is generated by a sequence of channels with some unknown and potentially channel dependent error, coherent errors can be averaged away into incoherent error. Indeed, some of these routines have been demonstrated experimentally \cite{Ware2018}, and been shown to reduce the coherence of noise. However, this benefit comes at the cost of additional compilation steps, which may in general increase circuit depth and require more sophisticated classical preprocessing.

@@

In this Letter, we take a different approach to address these problems at the gate synthesis level. We propose to inject additional decorrelating randomness into the system through the use of \emph{balanced control solutions} (BCSs). BCSs are families of control solutions that approximate the same target gate, but with balanced errors for any given instance of the noise Hamiltonian. That is, for a target gate, $U_T$, we seek a family of control solutions, $c_i(t)$, each implementing an approximation $U_i$ to the target gate, such that the family of unitary approximations is \emph{balanced}. A balanced family is one which satisfies, for some small $\alpha$,
\begin{equation}\label{eq:1}
  \frac{1}{N}\sum_{i=1}^N \omega_i U_i \rho U_i^\dagger = DPN[\alpha]\left(U_T \rho U_T^\dagger \right)
\end{equation}

@@

This techique can be used to turn coherent error into incoherent errorr, correlated error into uncorrelated error, and therefore non-Markovian noise into more Markovian noise. This is particularly, since many routines exist that can assess the quality of a gateset, however in the presence of non-Markovianity most of them become unreliable.

For example, randomized benchmarking and tomography will report incorrect answers without any syndrome \cite{Merkel2013}, and gateset tomography will report that the gateset failed to be Markovian, but will fail to diagnose in what way it was non-Markovian. Because BCSs can make error channels look more depolarizing, the correlations in the noise can be reduced, and the process can be made much closer to Markovian. In addition, we will show that this comes at no real cost to the fidelity of the implemented gate. Finally, unlike in \cite{Wallman2016, Campbell2017, Ware2018} this method of coherent error mitigation is cheap to implement - it only requires a random choice of pulse definition at run time, in addition to the extra storage required to store the pulse shapes - no classical preprocessing or runtime frame tracking is required.

@@

\section{Experimental Results}\label{experimental}
To demonstrate the realizability of this routine, we implemented it on a superconducting qubit, to calibrate an $RX(\frac{\pi}{2})$ pulse. The qubit being used had a FILL IN $\mu$s T1, and a FILL IN $\mu$s T2.

In this example we intentionally over and under-calibrated four pulseshapes. For each of these pulseshapes, in addition to their BCS and the calibrated pulse, we then ran a randomized benchmarking experiment.\cite{Magesan2011} The results are plotted in \ref{fig:rb}. In each case, our Clifford sequences were decomposed into RX($\frac{\pi}{2})$) and RY($\frac{\pi}{2})$) pulses.  In the superconducting qubit architecture used, these gates are implemented using the same pulse envelope definition. It was shown in \cite{Ball2016} that for a particular error model, non-Markovian, or \textit{quasi-static}, noise can result in more general Gamma Distributed noise, while Markovian noise, such as depolarizing noise, should result in Gaussian distributed fidelity estimates for each randomized benchmarking sequence depth. Our results are consistent with this, as demonstrated by the long tails in the four leftmost plots that vanish in the bottom right plot.

@@

\section{Conclusion and Future Work}
We have shown numerically and experimentally that drawing from a collection of implementations of a gate with the correct probabilties can reduce coherent error by an order of magnitude, while increasing incoherent error by a smaller amount, at virtually no cost to gate fidelity. We have demonstrated that these approximate controls can be generated through optimal control, and that the minimization problem is tractable. In addition, we have shown that it is possible to perform the routine on existing hardware. The code used to generate the numerical results in this paper is available at \cite{decorrelating_errors}.

@@

In addition, the numerical work in the paper assumed access to a model of the system. In general, an experimenter may not have a model readily available to describe the system, maybe due to on-chip crosstalk, or even if they do have access to such a model, it might be computationally intractable to simulate. In both of these situations, numerical generation of candidate pulseshapes and the minimization become infeasible. One approach would to use \textit{in situ} optimal control techniques \cite{Wu2018, Kelly2014, Ferrie2015} to generate candidate controls, and then use an optimizer like Nealder-Mead to minimize the magnitude of off-diagonal terms as measured by simple pulse sequences that amplify those elements of the process matrix.\needcite

>>>>

\newcommand{\erf}[1]{Eq.~(\ref{#1})}
\newcommand{\needcite}{{\color{blue}\textsuperscript{[citation needed]}}}
\newcommand{\note}[1]{{\color{red}[#1]}}
\newcommand{\kcy}[1]{{\color{red}[#1]_{\rm{KCY}}}}
\newcommand{\amp}[1]{{\color{red}[#1]_{\rm{AMP}}}}
%-------------Header begins here----------------------------------------
\begin{document}
% \tableofcontents
\title{Decorrelating Errors in Quantum Gates by Random Gate Synthesis}

@@

\begin{abstract}
Thresholds for fault-tolerant quantum computation are often calculated assuming a noise model in which errors are uncorrelated. While convenient for simulation, these error models are often unphysical. Work by Preskill and others has shown that arbitrarily long computations may be performed even in the presence of spatial correlation, provided the correlation is sufficiently weak and decays sufficiently quickly with distance, but at the cost of a significantly lower threshold. The success of algebraic decorrelation methods, such as dynamical decoupling, demonstrate that quantum control techniques are capable of reducing temporal noise correlations. We propose to introduce similar methods at the gate synthesis level to effect the $\amp{Why are we calling this spatial?}$spatial decorrelation of errors in quantum circuits, thereby increasing the threshold for fault-tolerant computation in such systems.
\end{abstract}

@@

Steady progress has been made in the theory of quantum error correction, proving higher thresholds for increasingly general models of noise \cite{Aharonov2006} $\kcy{more recent citations}$. These results show that quantum computation is feasible in principle, however recent NISQ \cite{Preskill2018} devices have noise that is not only often above thresholds, but that also violates fundamental assumptions made by the models used in these results, such as Markovianity \cite{Kitaev1997} and independence of errors\needcite. With these assumptions violated, promises about system performance and correctness are difficult to make.

Moreover, there are many ways in which device noise can fail to be Markovian, \cite{Kelly2018, BlumeKohout2017, Klimov2018}, and even when the noise \textit{is} Markovian, many models make simplifying assumptions about the structure of the noise. For instance, Pauli channels are often used to model systems\needcite due to their classical simulability\needcite, despite most noise sources being poorly approximated by Pauli channels.\needcite. Some authors have then taken the approach of using these easily-understood models to upper bound the error in a real system\cite{Puzzuoli2014}. This approach is correct and rigorous, but produces overly-pessimistic bounds on device performance.

Some characterization routines take a different approach, and assume that the noise is well-behaved. For example, randomized benchmarking and tomography will simple report incorrect answers without any syndrome in the case of non-Markovian noise \cite{Merkel2013}, and gateset tomography will report that the gateset failed to be Markovian.

Other authors have considering the problem of coherent error at the logical gate level, considering over rotations that might arise from gate approximation \cite{Campbell2017, 1612.01011}, and others have considered methods of \textit{twirling} error, such that it turns into a Pauli channel through random compilation methods\cite{Wallman2016, Ware2018}.

@@

In this Letter, we take a different approach to address the issues of coherent and non-Markovian error. We propose to inject additional decorrelating randomness into the system during gate synthesis through the use of \emph{balanced control solutions} (BCSs). BCSs are families of control solutions that approximate the same target gate, but with balanced errors for any given instance of the noise Hamiltonian. That is, for a target gate, $U_T$, we seek a family of control solutions, $c_i(t)$, each implementing an approximation $U_i$ to the target gate, such that the family of unitary approximations is \emph{balanced}. A balanced family is one which satisfies, for some small $\alpha$,
\begin{equation}\label{eq:1}
  \frac{1}{N}\sum_{i=1}^N \omega_i U_i \rho U_i^\dagger = DPN[\alpha]\left(U_T \rho U_T^\dagger \right)
\end{equation}

@@

Because BCSs produce error channels that are depolarizing, correlations in the noise are reduced, and by generating controls that produce BCSs over a range of parameter values, a system can be made less sensitive to non-Markovian errors, e.g. drift. In addition, we will give numerical results showing that the fidelity of the new approximation to $U$ is not much worse than the fidelities of the constituent controls.

@@

\section{Experimental Results}\label{experimental}
Here we present experimental results from implementing our routine on a superconducting transmon qubit. In this experiment we intentionally over and under-calibrated four pulses enacting approximate RX($\frac{\pi}{2})$) rotations. For each instatiation of the rotation, in addition to the BCS for the four pulses and a calibrated pulse, we ran a randomized benchmarking experiment.\cite{Magesan2011} The results are plotted in \ref{fig:rb}. In each case, our Clifford sequences were decomposed into RX($\frac{\pi}{2})$) and RY($\frac{\pi}{2})$) pulses. In the superconducting qubit architecture used, these gates are implemented using the same pulse envelope definition, and are therefore subject to the same miscalibration errors. It was shown in \cite{Ball2016} that for particular non-Markovian (\textit{quasi-static}) error models, noise will manifest as gamma distributed points for each sequence length. On the other hand, Markovian noise, such as depolarizing noise, will result in Gaussian distributed fidelity estimates for each randomized benchmarking sequence length. Our results are consistent with this $\amp{explain how, given new plot}$

@@

\section{Conclusion and Future Work}
We have shown both numerically and experimentally that drawing from a collection of different synthesized version of a gate with  optimized probabilties can reduce coherent error by at least an order of magnitude, at virtually no cost to gate fidelity. We have demonstrated that these approximate controls can be generated through optimal control, and that the minimization problem is tractable. In addition, we have shown that it is possible to perform the routine on existing hardware. We note that unlike in \cite{Wallman2016, Campbell2017, Ware2018} this method of coherent error mitigation is cheap to implement - it only requires a random choice of pulse definition at run time, in addition to the extra storage required to store the pulse definitions. In particular, no classical preprocessing or runtime frame tracking is required.

@@

In addition, the numerical work in the paper assumes access to a model of the system. In general, an experimentalist may not have a model readily available to describe the system, e.g. in the presence of unknown on-chip crosstalk. Even if they do have access to such a model, it might be computationally intractable to simulate. In both of these situations, numerical generation of candidate pulseshapes and the minimization become infeasible. One approach would then be to use \textit{in situ} optimal control techniques \cite{Wu2018, Kelly2014, Ferrie2015} to generate candidate controls, and then use an optimizer like Nealder-Mead to minimize the magnitude of off-diagonal terms as measured by simple pulse sequences that amplify those elements of the process matrix.\needcite The code used to generate the numerical results in this paper is available at \cite{decorrelating_errors}.
